### 2017-01-17 robust-software.tex
Shortened documentation rule

Numerous resources exist for writing high quality
documentation~\cite{karimzadeh2016} and so we do not cover it here, except for
two minimal elements: the README and usage.

The README is the first stop for most new users, usually available even before
the software is installed.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it. Usage is a terse, informative command-line help message that
guides the user in the correct use of the software.

Numerous guidelines exist on how to write good
READMEs\cite{Johnson1997,gnustandards};
their key common features are listed below. 
The README file for \withurl{khmer}{https://github.com/dib-lab/khmer/blob/master/README.rst}
is a good model to emulate.

\begin{description}

\item[\ruleminor{Explain what the software does.}] 
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

\item[\ruleminor{List required dependencies.}] Often, software depends on
specific versions of libraries, modules, or operating systems. Include the name and
version number for each dependency or a link to the requirements file. 
We address dependencies in more detail in Rule~5.

\item[\ruleminor{Provide compilation/installation instructions:}]
If the software needs to be compiled
or installed, list those instructions in the README.

\item[\ruleminor{List input and output files.}] All possible input and output files
should be listed in this section. 
If they use a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
If there is no rigorous format, as is common with log files, show at least a
few lines from an
example file and explain what the sections mean.

\emph{All} files should be listed, even those considered self-explanatory. Log
files are often full of valuable information that can be
mined for the user's specific purpose. If your users might need to know,
``Does this program report the percentage of reads trimmed to remove
adapter sequences?'' they should be able to check the README and confidently
say, ``Yes, it is in the log file''.

\item[\ruleminor{State attributions and licensing.}] Attributions are how you credit
your main contributors; licenses are how others may use and
credit your work.
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
\end{description}

In addition to a README, a robust program should
\ruleminor{print usage information when launching from the command line that explains the software's features}.
Usage information provides the first line of help for both first-time and
experienced users of command-line applications. Terseness is
important: usage that extends for multiple screens is 
difficult to read or refer to on the fly.

Usage is invoked either by running the software without
any arguments, running it with incorrect arguments, or by
explicitly choosing a help option.
More standard command-line behaviours are detailed in \cite{Seemann2013}.

Almost all command-line applications use a combination of
POSIX~\cite{posix2016} and GNU~\cite{gnustandards} standards for usage. 
The key features of a usage message are:

\begin{description}

\item[\ruleminor{The syntax for running the program}] This includes the
  name of the program and defines the relative location of optional
  and required flags, arguments and values for execution.  Arguments
  in {[}square brackets{]} are usually optional. An ellipsis (\ldots
  e.g. ``{[}OPTION{]}\ldots{}'') indicates that more than one value
  can be provided.

\item[\ruleminor{Primary function}] Similar to the README, the description
  reminds users of the software's primary function.

\item[\ruleminor{Most commonly used arguments, a description of each, and
    the default values}] Not all arguments need to appear in the
  usage, but those most commonly used should be listed. Users will
  rely on this for quick reference when working with the software.

\item[\ruleminor{Where to find more information}] Whether it's an email
  address, web site, or manual page, there should be an indication
  where the user can go to find out more.

\item[\ruleminor{Printed to standard output}] Usage should be displayed
  on standard output so that it can be piped into \texttt{less},
  searched with \texttt{grep}, or compared to the previous version
  with \texttt{diff}.

\item[\ruleminor{Exit with an appropriate exit code}] When usage is
  invoked by providing incorrect arguments, the program should exit
  with a non-zero code to indicate an error. However, when help is
  explicitly requested, the software should not exit with an error.

\end{description}

Documentation beyond the README and usage is up to the developer's discretion.
We think it is very important for developers to document their work, but our
experience is that people are unlikely do it during normal development, and we
don’t want to recommend practices that people won’t actually adopt. However, it
is worth noting that software that is widely used and contributed to has and
enforces the need for good documentation~\cite{gentleman2004}.

>>>>

How to write high quality documentation has been described elsewhere~\cite{karimzadeh2016}
and so here we only cover two minimal types: the README and usage. 
The README is usually available even before the software is installed, 
and exists
to get a new user started and point them towards more
help. Usage is a terse, informative command-line help message that
guides the user in the correct use of the software.

Numerous guidelines exist on how to \ruleminor{write a good
README file}~\cite{Johnson1997,gnustandards}.
Some key common features are listed below.

\begin{enumerate}

\item \textit{Explain what the software does}.
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

\item \textit{List required dependencies}.
We address dependencies in more detail in Rule~5.

\item \textit{Provide compilation or installation instructions}.

\item \textit{List all input and output files},
even those considered self-explanatory. 
Link to specifications for standard formats
and list the required fields and acceptable values in other files.
If there is no rigorous format, explain what the sections mean.
These files are often full of valuable information that can be
mined for the user's specific purpose.

\item \textit{List a few example commands} to get a user started quickly.

\item \textit{State attributions and licensing}. Attributions are how you credit
your contributors; licenses dictate how others may use and
need to credit your work.
\end{enumerate}

The program should also 
\ruleminor{print usage information} when launching from the command line.
Usage provides the first line of help for both new and experienced users. 
Terseness is important: usage that extends for multiple screens is 
difficult to read or refer to on the fly.

Almost all command-line applications use a combination of
POSIX~\cite{posix2016} and GNU~\cite{gnustandards} standards for usage. 
More standard command-line behaviours are detailed in \cite{Seemann2013}.
The key features of a usage message include:

\begin{enumerate}

\item \textit{Describe the syntax for running the program}, including the
  name of the program, the relative location of optional
  and required flags, other arguments, and values for execution.

\item \textit{Give a short description} to remind users of the software's primary function.

\item \textit{List the most commonly used arguments}, a description of each, and
    the default values.

\item \textit{State where to find more information}.

\end{enumerate}

Usage should be 
printed to standard output so that it can be 
combined with other bash utilities like \texttt{grep},
and it should finish with an appropiate exit code
depending upon whether it is a response to an error,
or was explicitly requested.

Documentation beyond the README and usage is up to the developer's discretion.
We think it is very important for developers to document their work, but our
experience is that people are unlikely do it during normal development
However, it
is worth noting that software that is widely used and contributed to has and
enforces the need for good documentation~\cite{gentleman2004}.
